There are no small Advantages: Reinforcement learning in near continuous time with Deep RelAtive Mixture of Advantages

### 2019-01-11 paper/icml_drau/advup.tex
Algorithm description.

However, the proposed reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$.
Indeed, for any given pair $V$, $A$,
\begin{align}
	\tilde{V}(s) &= V(s) - f(s)\\
	\tilde{A}(s, a) &= A(s, a) + \frac{f(s)}{\deltat}
\end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce, identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q^\pi_\deltat = Q$,
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
	A(s, a) = \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}

\subsection{Time invariant exploration}
\label{subsec:explo}
To obtain a time invariant reinforcement learning algorithm, a time invariant
exploration scheme is required.

The exploration scheme used~\cite{ddpg}, which consists in adding an
\emph{Ornstein-Uhlenbeck}~\cite{orn-uhl} process to actions, is naturally time
invariant in the sense that it is first defined as a time continuous stochastic
process, and that discretization of this process are used in practice. Indeed,
a continuous Ornstein-Uhlenbeck process is defined as
\begin{equation}
	dx_t = - (x_t - \mu) \theta dt + \sigma \dbt
	\label{eq:orn_uhl}
\end{equation}
and is naturally discretized by turning $d$'s into $\delta$'s, and using
the same $\deltat$ for both the learning algorithm and the exploration scheme.
Consequently, this exploration scheme can readily be used in continuous action
environments.

On the other hand, the usual exploration scheme used in discrete actions
Q-learning, $\varepsilon$-greedy exploration~\cite{sutton} is not time
invariant, in the sense that resulting state trajectories do not converge to a
well defined stochastic process as $\deltat \rightarrow 0$. As $\dt$ goes to $0$,
for any $\varepsilon$ independant of $\dt$, the behavior of state trajectories get
closer to that of a white noise.

To obtain a temporally coherent exploration scheme in discrete action environment
when using Q-learning we instead directly apply a discretization of a time continuous
stochastic process to the output of the approximate Q-function, and select actions
by maximizing the resulting perturbed Q-function. Typically, the approximate Q-function
values are also perturbed using a discretized Ornstein-Uhlenbeck noise.

\subsection{Algorithms}
\label{subsec:algorithm}
Two variants of \emph{Deep Advantage Updating} are derived to cover both the continuous
and discrete action cases. They both rely on variants of Q-learning. More precisely,
we learn $V$ and $A$ using approximate version of the following bellman equation
\begin{align}
	V^{\pi}(s) + A^{\pi}(s, a) &= \E_{\delta r \sim \delta r(s, a)}\delta r\nonumber\\ 
				   &+ \gamma^{\deltat}  \E_{\delta s \sim \delta s(s, a)} V^{\pi}(s + \delta s)\label{eq:bellman_A}\\
	A^{\pi}(s, \pi(s)) &= 0\label{eq:max_A}
\end{align}
for a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi(s, a')$.
Eq.~\eqref{eq:max_A} is directly verified owing to the parameterization of $A$.
To approximately verify~\eqref{eq:bellman_A}, the corresponding squared residual
is minimized by an approximate gradient descent.
The corresponding update equations when learning from incoming transitions of
an exploratory trajectory are
\begin{align}
	\delta Q_t &= A(s_t, a_t) - \frac{\delta r_t + \gamma^{\deltat} V(s_{t+\deltat}) - V(s_t)}{\deltat}\\
	\theta^1_{t+\deltat} &= \theta^1_t + \eta_1 \partial_{\theta^1} V(s_t) \delta Q_t\\
	\theta^2_{t+\deltat} &= \theta^2_t + \eta_2 \partial_{\theta^2} A(s_t, a_t) \delta Q_t.
	\label{eq:approx_bellman_A}
\end{align}
where $\eta$'s are learning rates. We are now left to derive appropriate learning rates
to provide invariance to time discretization.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
\end{algorithm}

>>>>

However, the proposed reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$.
Indeed, for any given pair $V$, $A$,
\begin{align}
	\tilde{V}(s) &= V(s) - f(s)\\
	\tilde{A}(s, a) &= A(s, a) + \frac{f(s)}{\deltat}
\end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce, identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q^\pi_\deltat = Q$,
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
	A(s, a) = \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi_\deltat$,
from which a proper hiearchy of actions can be derived for any $\deltat$.

\subsection{Time invariant exploration}
\label{subsec:explo}
To obtain a time invariant reinforcement learning algorithm, a time invariant
exploration scheme is required. All the exploration schemes hereby considered
generate sequences of actions as
\begin{equation}
	a^k_\deltat = \pi^\text{explore}_\deltat(s^k_\deltat, \nu^k_\deltat)
\end{equation}
with $(\nu^k_\deltat)_k$ a sequence of random variables, independant from the $a$'s and $s$'s.
For instance, $\varepsilon$-greedy exploration naturally falls in this framework, with each $\nu^k_\deltat$
being a pair composed of a bernoulli random variables $b^k_\deltat$ with probability of being $1$ equal to $\varepsilon$,
and of a uniform categorical random variable $\xi^k_\deltat$ on actions. Exploratory actions are then selected as
\begin{equation}
	a^k_\deltat = (1 - b^k_\deltat) \text{argmax}_{a'} Q^\pi_\deltat(s, a') + b^k_\deltat \xi^k_\deltat.
\end{equation}

A sufficient condition for the exploratory policy to admit a time continuous
limit is for $\pi^\text{explore}_\deltat$ to be a measurable function, and for
$\nu_\deltat$ to admit one, i.e. that the sequence $\nu_\deltat$ converges to a
well defined continuous stochastic process $\nu$ as $\deltat$ goes to $0$.
\TODO{I'm convinced that this is true, reviewers might not be.}

The exploration scheme used~\cite{ddpg}, which consists in adding an
\emph{Ornstein-Uhlenbeck}~\cite{orn-uhl} process (OU process) to actions, is naturally time.
invariant. It can be rewritten as
\begin{equation}
	a^k_\deltat = \pi(s^k_\deltat) + \nu^k_\deltat
\end{equation}
where $\nu^k_\deltat$ is the discretization of a time continuous \emph{Ornstein-Uhlenbeck}
process, defined as
\begin{equation}
	d\nu_t = - \nu_t \theta dt + \sigma \dbt.
	\label{eq:orn_uhl}
\end{equation}

On the other hand, the same argument cannot be applied to $\varepsilon$-greedy
exploration, as the $\nu_\deltat$'s used to define it do not converge to a 
well defined stochastic process, but to a white noise.\footnote{\TODO{$\varepsilon$-greedy
		does not, in general, converges to a well defined exploration scheme in continuous time.
		For instance, if actions directly affects position, $\varepsilon$-greedy results in white-noise
trajectories in state space. Show this in more detail in appendix ?}}

To obtain a temporally coherent exploration scheme in discrete action environment
$\nu_\deltat$ is chosen to be a discretization of a $|\mathcal{A}|$ dimensional
continuous OU process, and set
\begin{equation}
	a^k_\deltat = \text{argmax}_{a'} \left(Q^\pi_\deltat(s, a') + \nu_\deltat^k[a']\right)
\end{equation}
where $\nu_\deltat^k[a']$ denotes the $a'$-th component of the discretized OU process at time
$k$. The resulting exploration scheme converges to a proper time continuous exploration scheme.

\subsection{Algorithms}
\label{subsec:algorithm}
To learn $V_{\theta_1}$ and $A_{\theta_2}$, variants of Q-learning for continuous~\cite{ddpg} and
discrete~\cite{dqn} action spaces are used. In details, $V_{\theta_1}$ and $A_{\theta_2}$ are learnt
using approximate version of the following bellman equation
\begin{align}
	V^{\pi}(s) + A^{\pi}(s, a) &= \E_{\delta r \sim \delta r(s, a)}\delta r\nonumber\\ 
				   &+ \gamma^{\deltat}  \E_{\delta s \sim \delta s(s, a)} V^{\pi}(s + \delta s)\label{eq:bellman_A}\\
	A^{\pi}(s, \pi(s)) &= 0\label{eq:max_A}
\end{align}
for a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi(s, a')$,
exactly implemented in the case of discrete actions, and approximated by a
neural network $\pi_\phi(s)$, trained to maximize (an approximation of)
$A^\pi(s, \pi_\phi(s))$ in the continuous case.  Eq.~\eqref{eq:max_A} is
directly verified by $A_{\theta_1}$, owing to the reparameterization described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The corresponding update equations when learning from incoming transitions of
an exploratory trajectory are
\begin{align}
	\delta Q_t &= A(s_t, a_t) - \frac{\delta r_t + \gamma^{\deltat} V(s_{t}+\delta s_{t}) - V(s_t)}{\deltat}\\
	\theta^1_{t+\deltat} &= \theta^1_t + \eta_1 \partial_{\theta^1} V(s_t) \delta Q_t\\
	\theta^2_{t+\deltat} &= \theta^2_t + \eta_2 \partial_{\theta^2} A(s_t, a_t) \delta Q_t.
	\label{eq:approx_bellman_A}
\end{align}
where $\eta$'s are learning rates. In practice, $A$ and $V$ are learned using
transitions sampled from a replay buffer instead of trajectory samples, to
prevent excessive correlation between training samples~\cite{dqn}. Pseudocode
for the discrete actions algorithm is given in Alg.~\ref{alg:dau}. Pseudocode
for the continuous actions algorithm is given in the supplementary material.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}
